---
title: "Problem Set 3"
author: "Connor Tragesser, Discussants are noted below"
date: "2023-09-10"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---
Discussed with Abdullah Jaber, Kristen Adams, Edward Howe, Jinglong Guo, Jae Chang, Jae Eun Jun, Soyun Chang, Ryo Sawayma, Cece Kinney, and Alice Moon

Begin by preparing the environment. The working directory needs to be set where your data will be available (meaning, you will need to change this for your computer), and the packages that are necessary for this exercise need to be loaded from the library (this assumes they are already installed). I'm also clearing the environment before beginning.

```{r setup}
library(foreign)
library(tidyverse)
library(modelsummary)
library(formatR)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(root.dir = "/Users/connortragesser/Documents/Political Science Coursework Readings/Quant 1/Problem Set 3/")

getwd()
rm(list = ls()) 

```

## Problem 1

### Part A

The answer is "1. 35/36".

### Part B

The answer is "3. 53". The question tells us being left handed is independent of political preference, so we would still expect that a left handed person has a .53 chance of voting for Obama. Thought of mathematically, given the formula from the notes, for independence, $P(B|A) = P(B)$ in this case, $P(vote\;for\;obama |left\;handed) = P(vote\;for\;obama)$.

## Problem 2

### Part A

```{r}
124/164
```

There is a 75.6% chance the country has a per capita GDP below \$5000.

### Part B

```{r}
(64+29)/164
```

There is a 56.7% chance the country has a proportional representation system.

### Part C

$P(PR\;system|GDP\;above\;5000) = .725$

```{r}
29/40
```

There is a 72.5% chance the country has a proportional representation system given that it has a Per Capita GDP \> \$5000.

### Part D

```{r}
64/(64+29)
```

There is a 68.8% chance that the country has a pcGDP below \$5000 given that it has a PR system.

### Part E

Having a PCGDP above \$5000 is not independent of using a PR system. If two events are independent, $P(B|A) = P(B)$. This is not the case, given that, in our scenario, $P(GDP\;above\;5000) = 1 - .756 =.244$, while $P(GDP\;above\;5000|PR\;system) = \frac{P(GDP\;above\;5000)\times P(PR\;system|GDP\;above\;5000)}{P(PR\;system)} = .312$. These values are unequal.

```{r}
1-.756

(.244*.725)/.567
```

## Problem 3

The information contained in the problem, along with the needed properties, are listed below. Where $U = Support$ and $Z = Says\;Support$

$$P(Z|U) = 1$$ $$P(Z) = 0.85$$ $$P(Z|U^c) = 0.2$$ $$P(Z^c) = 0.15$$ $$(1)\;P(A) = P(A|B)P(B)+P(B|A^c)P(B^c)$$ $$(2)\;P(A) + P(A^c) = 1$$

We are interested in solving for $P(U)$ so we can use the given value we have for $P(Z)$ in equation $(1)$:

$$P(Z) = P(Z|U)P(U)+P(Z|U^c)P(U^c)$$

Plugging in the known values we have, the following is given:

$$0.85 = 1 \times P(U) + 0.2 \times P(U^c)$$ Using $(2)$ with some algebra applied, and dropping the multiplication of 1, this can be rewritten as

$$0.85 = P(U) + 0.2 \times (1-P(U))$$

Which can be distributed to

$$0.85 = P(U) + 0.2 - 0.2(P(U))$$

Doing simple algebra and combination of like terms we get

$$.65 = .8(P(U))$$

And

$$.8125 = P(U)$$

Thus, there is a 81.25% that a randomly selected respondent actually supports the ruling party.

## Problem 4

### Part A

We cannot know based on the given information, because we don't know the total distribution of men and women.

### Part B

This is true.

### Part C

This is true.

### Part D

This is true.

### Part E

This is true.

## Problem 5

```{r}

qog <- read.dta("QoG_2010.dta")
head(qog)

```

### Part A

```{r}
boxplot(qog$undp_gdp, xlab = "per capita GDP ($)", ylab = "Amount")
```

This boxplot conforms with my expectations, wherein a large group of countries are bunched in the lower end, and there are a few outliers at the very high end of the distribution. I would expect the mean to be grater than the median, because the outliers on the high end would push the mean up, but the median would remain unaffected.

### Part B

```{r}
ggplot(qog, aes(x=fh_status, y=undp_gini)) + 
    geom_boxplot() + 
  xlab("Freedom Index") +
  ylab("Gini Coefficient") +
  theme_bw()
```

This graph shows that, among the three levels of freedom, "Partly Free" has the highest median Gini coefficient, followed by "Not Free" and then "Free". The IQR for the levels are relatively similar, with the "Not Free" one being the closest together. The "Free" category seems to have the widest range of Gini coefficients. Substantively, because Gini coefficients rise as the inequality rises, this indicates that "Free" countries, as defined here, have the lower median amount of inequality than the other two, and "Partly Free" countries tend to be the most unequal, both in terms of median and higher end of the IQR.

### Part C

```{r}
ggplot(qog, aes(x = undp_gdp, y = wdi_gris)) +
  geom_point() +
  theme_bw() +
  xlab("GDP Per Capita") +
  ylab("% Girls to Boys Enrolled in Schools")
```

This relationship is difficult to parse because of the outliers in GDP per capita. However, there seems to be a positive relationship between these two variables. No relatively rich countries have very low rates of girls in schools, and past a relatively low level of per capita GDP, there is generally parity between girls and boys in school enrollment. This relationship may come about because high education in general may lead to a high per capita GDP, and countries with higher rates of girls in school probably have a higher education rate (it is unlikely that many countries have a noticeably different number of girls and boys). It could also be the case that as countries grow in economic sophistication, thus increasing per captia GDP, they may develop more liberal social values as a result of market forces, making the populous more likely to support education for girls.
