---
title: "Problem Set 9"
author: "Connor Tragesser"
date: "2023-10-28"
output: pdf_document
---

Discussed with Alice Moon, Ryo Sawayama, SoYun Chang, Jae Eun Jun, Jinglong Guo, Edward Howe, and Jae Chang. 

Begin by preparing the environment. The working directory needs to be set where your data will be available (meaning, you will need to change this for your computer), and the packages that are necessary for this exercise need to be loaded from the library (this assumes they are already installed). I'm also clearing the environment before beginning.

```{r setup}
library(foreign)
library(tidyverse)
library(modelsummary)
library(formatR)
library(data.table)
library(readxl)
library(gmodels)
library(gtsummary)
library(gt)
library(flextable)
library(kableExtra)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(root.dir = "/Users/connortragesser/Documents/Political Science Coursework Readings/Quant 1/Problem Set 9/")

getwd()
rm(list = ls()) 

set.seed(2453)

```

## Problem 1

### Part A

The expectation is $$\frac{row\;total \times column\;total}{total\;sample\;size}$$

### Part B

The marginal distribution is simply (200,200).

## Problem 2

### Gailmard 8.4

We have a an observation of 5 taken from a poisson distribution. We want to test whether the poisson distribution it was taken from has $\lambda = 3$.  

$$H_0: \lambda = 3$$
$$H_1: \lambda > 3$$

$$observed = 5$$

```{r}
ppois(5, lambda = 3, lower.tail = FALSE)
```

We fail to reject the null hypothesis that the distribution has $\lambda = 3$. 

### Gailmard 8.5

Develop DGP first

```{r}
lamda_dist <- rpois(40, lambda = 5)
```

### Part A
```{r}
t.test(lamda_dist, mu = 5)
```
This test fails to reject the null hypothesis that the mean of the sample distribution is equal to 5 at the $\alpha = 0.05$ level.


### Part B
```{r}
t.test(lamda_dist, mu = 8)
```
This test rejects the null hypothesis that the mean of the sample distribution is equal to 8 at the $\alpha = 0.05$ level.

## Problem 3

Load dataset 

```{r}
culture <- read.dta("apsrfinaldata.dta")
head(culture)
```


### Part A

The table shows that a one unit increase in between gorup inequality is associated with an average decrease of .228 in public goods, holding the standardized values of the natural logarithm of GDP per capita equal. Similarly it shows that a one unit increase in the standardized values of the natural logarithm of GDP per capita is associated with an average increase of .800 in public goods holding between gorup inequality equal. 

```{r}
reg_model_culture <- lm(pg ~ betweenstd + lngdpstd, data = culture)
summary(reg_model_culture)
modelsummary(reg_model_culture, stars = TRUE)
```

### Part B

The standard error for the betweenstd coefficient is .075 and the standard error for the lngdpstd coefficient is .075. Some rounding is done here.

### Part C

The reported t-statistics are calculated as a test for whether the observed coefficients are significantly different from 0. Under the bull hypothesis, the true beta coefficient that we do not observe equals 0, or $H_0: \beta = 0$, while our alternative hypothesis is $H_0: \beta \neq 0$. The test statistic is calculated like so, for each coefficient respectively:

$$T = \frac{\hat\beta - 0}{\hat {se}[\hat\beta_1]}$$

The t-stat for betweenstd is -3.036 and its p-value is 0.00406. This means that we reject the null hypothesis that $\beta = 0$, based on the observed beta coefficient, $\hat\beta$, and its standard error. For lngdpstd, the t-stat is 10.633 and the p-value is .00000000000013. This also means that we reject the null hypothesis that $\beta = 0$, based on the observed beta coefficient, $\hat\beta$, and its standard error.

## Problem 4

Loading the data 

```{r}
mar <- read.dta("MAR_2006.dta")
head(mar)
```

### Part A

```{r}

mar <- mar %>% 
  mutate(POLGR_str = ifelse(POLGR == 0, str_c("No political", "grievances expressed", sep = "\n"), 
                            ifelse(POLGR == 1, str_c("Political grievances focused on", "ending discrimination", sep = "\n"),
                                   ifelse(POLGR == 2, str_c("Political grievances focused on", "creating or strengthening", "remedial policies", sep = "\n"),
                                          ifelse(POLGR == 3, str_c("Political grievances focused on", "creating or strengthening", "autonomous status", sep = "\n"),
ifelse(POLGR == 4, str_c("Political grievances focused on", "creating separate state for group", "or revanchist change in borders", sep = "\n"), "")))))) %>% 
  mutate(EXECREP_str = ifelse(EXECREP == 0, str_c("No group representation", "in central government", sep = "\n"), str_c("Group representation", "in central government", sep = "\n"))) %>% 
  mutate(GROUPCON_str = ifelse(GROUPCON == 0, "Widely dispersed", 
                                ifelse(GROUPCON == 1, "Primarily urban or minority in one region",
                                       ifelse(GROUPCON == 2, "Majority in one region, others dispersed",
                                              ifelse(GROUPCON == 3, "Concentrated in one region", "")))))


mar$POLGR_str <- factor(mar$POLGR_str, levels = unique(mar$POLGR_str[order(mar$POLGR)]), ordered = TRUE)
mar$EXECREP_str <- factor(mar$EXECREP_str, levels = unique(mar$EXECREP_str[order(mar$EXECREP)]), ordered = TRUE)
mar$GROUPCON_str <- factor(mar$GROUPCON_str, levels = unique(mar$GROUPCON_str[order(mar$GROUPCON)]), ordered = TRUE)


bar_chart_maker <- function(var, title, xlabel) {
  ggplot(aes(x = {{var}}), data = mar) +
    geom_bar() +
    theme_bw() +
    ggtitle(title) +
    scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
    xlab(xlabel) +
    ylab("Count")
}

bar_chart_maker(POLGR_str, "Group Political Grievance", "Group Political Grievance")
bar_chart_maker(EXECREP_str, "Group Representation in Executive Branch of Central Government", "Group Representation in Executive Branch of Central Government")
bar_chart_maker(GROUPCON_str, "Group Geographical Concentration", "Group Geographical Concentration")

```

### Part B

The chi squared test allows us to test whether the distributions of subgroups match expectations based on the whole, as in $\frac{row\;total \times column\;total}{table\;total}$. Effectively, it is a test of independence between two variables. The chi squared test allows us to reject the null hypothesis of independence at the $\alpha = .05$ level. Given the clean empirical frequency/proportion table, this makes sense. At the total level, 59% of observations have "No group representation in central government" while 41% of observations have "Group representation in central government". When looking at some of the subgroups of political grievance, however, this distributions change drastically, from 45% and 55% for the lowest level of conflict, to 73% and 27% for the highest. The dirty theoretical table is the expected number of cases found in each political representation group if the two variables in question were completely independent As you can see, it looks very different from the empirical table. 

```{r}

chisq.test(x = mar$EXECREP, y = mar$POLGR)

freq_table <- mar %>% 
  select(POLGR_str, EXECREP_str) %>% 
  tbl_cross(percent = "row", label = list(POLGR_str ~ "Group Political Grievance",
                                             EXECREP_str ~ "Group Rep in Executive Branch")) %>%
  as_flex_table(width = 10)
freq_table

props <- mar %>% 
  group_by(EXECREP) %>% 
  summarise(num = n()) %>% 
  mutate(prop = num/sum(num)) %>% 
  select(EXECREP, prop) %>% 
  spread(EXECREP, prop) 

exp_table <- mar %>% 
  group_by(POLGR_str) %>% 
  summarise(total_no_rep = n()) %>% 
  mutate(total_with_rep = total_no_rep)

exp_table$total_no_rep = exp_table$total_no_rep * props$`0`
exp_table$total_with_rep = exp_table$total_with_rep * props$`1`
exp_table$total = exp_table$total_no_rep + exp_table$total_with_rep
exp_table

```





