---
title: "Problem Set 10"
author: "Connor Tragesser"
date: "2023-11-09"
output: pdf_document
---

Discussed with Edward Howe, Kristen Adams, SoYun Chang, Abdullah Jaber, Jae Eun Jun, Jinglong Guo, Cece Kenney, and Jae Chang.

Begin by preparing the environment. The working directory needs to be set where your data will be available (meaning, you will need to change this for your computer), and the packages that are necessary for this exercise need to be loaded from the library (this assumes they are already installed). I'm also clearing the environment before beginning.

```{r setup}
library(foreign)
library(tidyverse)
library(modelsummary)
library(formatR)
library(data.table)
library(readxl)
library(scales)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(root.dir = "/Users/connortragesser/Documents/Political Science Coursework Readings/Quant 1/Problem Set 10/")

getwd()
rm(list = ls()) 

set.seed(2453)

```

## Gailmard 9.1

### Part A

We need to use a hypothesis test for proportions. The null hypothesis is that $\pi = .5$. We observed $\pi = .51$

```{r}

z = (.51-.5) / ((.5*.5) / 625)^(1/2)
2 * (1 - pnorm(z))
```

This test fails to reject the null hypothesis 50% of the population (American adults) believe that economic growth should have a higher priority than environmental protection at the $\alpha = 0.05$ level.

### Part B

No I did not need to bootstrap the standard error. Bootstrapping is used when the theoretical distribution is unknown or too complicated to make assumptions about. Here, the distribution of differences can be assumed to be standard normal. 

### Part C

Confidence intervals are calculated using Z-scores. At the 95% level, it would be in the range $-1.96 \le z \le 1.96$, finding the values that get those z-scores, and calculating the distance. 

$$-1.96 = \frac{H_1 - .51}{\sqrt{\frac{.51(1-.51)}{625}}}$$
Therefore,

$$H_1 = -1.96 \times \sqrt{\frac{.51(1-.51)}{625}} - .51$$
And the other direction would replace $-1.96$ with $1.96$

```{r}

lower_bound = -1.96 * ((.51*(1-.51)) / 625)^(1/2) - .51
upper_bound =  1.96 * ((.51*(1-.51)) / 625)^(1/2) - .51

moe = ((upper_bound - lower_bound)/2)*100

moe
```

The margin of error for this poll is 3.92%

## Gailmard 9.2

Putting some proportions together first

```{r}
prop_men_dem = 90/220
n_men = 220
prop_women_dem = 145/243
n_women = 243
n_tot = 220 + 243
tot_prop = (90 + 145) / (220 + 243)
```


### Part A

Using critical value of 1.96 to get the 5% level:
```{r}
lower_bound = -1.96 * sqrt((tot_prop * (1 - tot_prop))/ n_tot)
upper_bound = 1.96 * sqrt((tot_prop * (1 - tot_prop))/ n_tot)

paste("The CI is ", 100 * round(tot_prop + lower_bound, 2), "% to ", 100 *round(tot_prop + upper_bound, 2), "%.", sep = "")
```


### Part B

$H_0$: Prop females who are dem = .5 and $H_0$: Prop males who are dem = .5  for the two tests, respectively. The alternative is that $H_1$: Prop females who are dem $\ne$ .5 and $H_0$: Prop males who are dem $\ne$ .5 for the two tests. 

$$SE_\pi = \sqrt{\frac{\pi(1-\pi)}{N}}$$
$$z = \frac{\hat\pi_1 - \pi_0}{SE_\pi}$$ 

```{r}
z_women = (prop_women_dem-.5) / ((.5*.5) / n_women)^(1/2)
z_men = (prop_men_dem-.5) / ((.5*.5) / n_men)^(1/2)

2 * pnorm(z_men)
2 * (1 - pnorm(z_women))
```

In both tests, we reject the mull hypothesis that the group is equally likely to identify as Republican and Democrat.

### Part C

pooled 2 sample proportion test

$$z = \frac{\hat\pi_{man\;dem} - \hat\pi_{woman\;dem}}{SE}$$ 

where,
$$SE = \sqrt{\hat\pi \times (1 - \hat\pi)\times(\frac{1}{n_{men}}+\frac{1}{n_{women}}))}$$

where $\hat\pi$ is the total proportion of the 2 samples. 

Null Hypothesis: $$H_0: \hat{\pi}_{man\;dem} - \hat{\pi}_{woman\;dem} = 0$$
Alternative: $$H_1: \hat{\pi}_{man\;dem} - \hat{\pi}_{woman\;dem} \neq 0$$
```{r}
se = sqrt(tot_prop * (1 - tot_prop) * ((1/n_men) + (1/n_women)))
z_pooled_prop = ((prop_men_dem - prop_women_dem))/se
2 * pnorm(z_pooled_prop)
```


This test allows us to reject the null hypothesis that the two proportions, % of men who were Democrats and % of women who were Democrats, are the same. 

## Gailmard 9.3

It is *not* true that two variables that have overlapping 95% confidence intervals are not significantly different. To say that two variables are not significantly different, the *mean* of each variable has to fall within the the 95% confidence interval of the other. If this is not the case, it is possible that the 95% confidence interval for their difference does not include 0.  

## Gailmard 9.4

The formula for the bounds of a one-sided confidence interval at the 95% level is 

$$-\inf \le \mu \le \hat \theta + 1.65(\frac{\sigma}{\sqrt N})$$

where $\mu$ is the expected value of $\hat \theta$ and $\sigma$ is the standard deviation. It is wider than the 2 sided test at the same level of confidence because we have an area of 5% of that right hand side of testing, rathern than 2.5% (which is 5% split in two directions). 



